Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
Fixing typos. Explaining diff
more typos
Acked-by: Julian Bangert <jbangert@acm.org>
xx note
nit
pass through 5.1

@@

Transformations can handle a wide variety of patterns in data formats, including the following: 
\paragraph{Offsets.}
We provide a transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$file\_stream, \@offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}.

@@

\paragraph{}
A real-world example with many different transforms, used to support
the Zip file format, is described in \S\ref{s:eval-format-zip}.

>>>>

\XXX[nz][Do you need \cc{*} here?  It seems like \cc{optional paren} would
already have a C representation of \cc{paren*} according to our big figure.]

@@

Transformations can handle a wide variety of patterns in data formats, including the following: 
\paragraph{Offsets.}
We provide a transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
be used to parse data at the offset contained in \cc{@offset}, by
using \cc{apply \$fragment some_parser}.

@@

\paragraph{}
A real-world example with many different transforms, used to support
the ZIP file format, is described in \S\ref{s:eval-formats}.

### 2014-05-01 paper/osdi14/eval.tex
pass through 5.1

\input{fig-zip}
\label{s:eval-format-zip}
Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf:berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

Duplicating information made sense when ZIP files were
stored on floppy disks with slow seek times and high fault rates and memory constraints made it
impossible to keep the ZIP directory in memory or the archive was split across multiple disks.
However, care must be taken that the meta-data is consistent. For example, vulnerabilities could occur
if the length in the central directory is used to allocate memory and the length in the local
directory is used to extract without checking that they are equal first.

We have implemented a grammar for ZIP. A strongly abbreviated version is shown in
Figure~\ref{fig:zip-extract}.

Our ZIP grammar is a good example of transformations can capture complicated syntax in real-world
file format. The \cc{zip\_file} grammar first splits the entire file stream into two streams
based on the \cc{zip\_end\_of\_directory} transform.
\cc{zip\_end\_of\_directory_parse(NailStream *files, NailStream *dir, NailStream (current))}
finds the end of directory header as described above by scanning the file backwards and splits the
file into two streams, one containing the end-of-directory header and one containing the file
contents. The header stream is then parsed. The file stream, to which all the offsets in the
metadata refer, is passed as an argument to the \cc{end\_of\_directory} parser.
The directory header is then parsed and the offset and size transforms provided with Nail are used
to isolate the actual directory from the file stream. Each directory entry in turn points to a local
file header.

The \cc{file} grammar for a zip file entry is particularly interesting. It receives the file
size, checksum and compression method of the file from the directory header. However, this same
information is duplicated in the file entry, so the grammar uses a standard \cc{depend}
transform to check whether the two values are equal. Unlike most other transforms, \cc{depend}
does not consume or produce strings, it only checks that two dependent fields are equal when parsing
and assigns the value of the second field to the first when generating. Immediately following the
file entry is the compressed data. Because most compression algorithms operate on unbounded streams
of data, Nail decompresses data in two steps.
First, it isolates the compressed data from the rest of the stream by using the size transform. In
this case, size operates on the current stream, which means it will consume data starting at the
current position of the parser in the input.
Second, Nail invokes a custom transform that implements the appropriate compression and decompression
functions based on the specified compression method. These functions are otherwise oblivious about
the layout or metadata of the file.

Existing parser languages are not able to handle a file format of this complexity.

\noindent The results suggest that Nail is a good fit for these data formats.

>>>>

An especially tricky data format is the ZIP compressed archive
format~\cite{pkzip}.  ZIP files are normally parsed end-to-beginning. At
the end of each ZIP file is an \emph{end-of-directory header}. This header
contains a variable-length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid
length field is found. Many ZIP implementations disagree on how to find
this header in confusing situations, such as when the comment contains
the magic number~\cite{wolf:berlinsides-zip}.  This end-of-directory
header contains the offset and size of the \emph{ZIP directory}, which
is an array of \emph{directory entry headers}, one for every file
in the archive.  Each entry stores file metadata, such as file name,
compressed and uncompressed size, and a checksum, in addition to the
offset of a \emph{local file header}. The local file header duplicates
most information from the directory entry header and the compressed file
contents follow it immediately.

\input{fig-zip}

Duplicating information made sense when ZIP files were stored on floppy
disks with slow seek times and high fault rates, and memory constraints
made it impossible to keep the ZIP directory in memory or the archive
was split across multiple disks.  However, care must be taken that the
metadata is consistent. For example, vulnerabilities could occur if
the length in the central directory is used to allocate memory and the
length in the local directory is used to extract without checking that
they are equal first, as was the case in the Python ZIP
library~\cite{cve-python-zip-maybe}.

\begin{figure}[tb]
\smaller[0.5]
\input{code/zip-eod-xform}
\caption{Signatures of stream transform functions for handling the
end-to-beginning structure of ZIP files.}
\label{fig:zip-eod-xform}
\end{figure}

Figure~\ref{fig:zip-extract} shows an abbreviated version of our ZIP
file grammar.  The ZIP grammar is a good example of how transformations
capture complicated syntax in a real-world file format; existing parser
languages cannot handle a file format of this complexity.

The \cc{zip\_file} grammar first splits the entire file stream into two
streams based on the \cc{zip\_end\_of\_directory} transform (whose two
C function signatures are shown in Figure~\ref{fig:zip-eod-xform}).
\cc{zip\_end\_of\_directory_parse} finds the end-of-directory header
as described above, by scanning the file backwards, and splits the
file into two streams, one containing the end-of-directory header and
one containing the file contents.  The header stream is then parsed,
according to the \cc{apply} rule, using the \cc{end\_of\_directory}
parser.  The file stream, to which all the offsets in the metadata refer,
is passed as an argument to that parser.  The directory header is then
parsed, and the offset and size transforms provided by Nail are used
to isolate the actual directory from the file stream.  Each directory
entry in turn points to a local file header.

The \cc{file} grammar for a ZIP file entry is particularly interesting.
It receives as arguments the file size, checksum, and compression method
of the file from the directory header.  However, this same information
is duplicated in the file entry, so the grammar uses the Nail-supplied
\cc{depend} transform to check whether the two values are equal. Unlike
most other transforms, \cc{depend} does not consume or produce strings; it
only checks that two dependent fields are equal when parsing, and assigns
the value of the second field to the first when generating.  This ensures
that the programmer does not have to worry about inconsistencies when
handling the internal representation of a ZIP file.

Immediately following the file entry is the compressed data.  Because
most compression algorithms operate on unbounded streams of data, Nail
decompresses data in two steps.  First, it isolates the compressed
data from the rest of the stream by using the \cc{size} transform,
which operates on the current stream, meaning it will consume data
starting at the current position of the parser in the input.  Second,
Nail invokes a custom \cc{zip\_compression} transform that implements
the appropriate compression and decompression functions based on the
specified compression method. These functions are otherwise oblivious
to the layout or metadata of the file.

### 2014-05-01 paper/osdi14/intro.tex
pass through 5.1

Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
existing parser generators. Those parsers cannot directly deal with length or checksum fields
directly, so the programmer has to either write potentially unsafe code to deal with such features,
or build contrived grammar constructs, such as introducing one grammar rule for each possible value
of a length field. Offset fields usually require the programmer to manipulate a parser's internal
state to re-position its input. More complicated transformations, such as
handling compressed data, cannot be implemented at all.

@@

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for Zip files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to parse and generate external
data representations.

>>>>

Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
existing parser generators. Those parsers cannot directly deal with length or checksum fields
directly, so the programmer has to either write potentially unsafe code to deal with such features,
or build contrived grammar constructs, such as introducing one grammar rule for each possible value
of a length field. Offset fields usually require the programmer to manipulate a parser's internal
state to re-position its input. More complicated transformations, such as
handling compressed data, cannot be represented at all.

@@

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for ZIP files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to parse and generate external
data representations.

### 2014-05-01 paper/osdi14/relwk.tex
pass through 5.1

We presented an earlier design of Nail at a
workshop~\cite{bangert:nail-langsec-anon}.  At that stage, Nail had only
limited support for dependent fields, and did not support stream
transforms at all, which are crucial for supporting real-world formats
like DNS and Zip.  The workshop paper also did not provide a detailed
design discussion or evaluation.

>>>>

We presented an earlier design of Nail at a
workshop~\cite{bangert:nail-langsec-anon}.  At that stage, Nail had only
limited support for dependent fields, and did not support stream
transforms at all, which are crucial for supporting real-world formats
like DNS and ZIP\@.  The workshop paper also did not provide a detailed
design discussion or evaluation.
